The Impact of Accretion Disk Winds on the Optical Spectra of Cataclysmic Variables

### 2015-04-08 draft_mnras.tex
trying to make final submitted zip
fiddled with orders. ArXix? would want email address

%%%% Various options for document class.
%\documentclass[usenatbib, a4paper]{preprint}
\documentclass[preprint, a4paper, 11pt]{aastex}
%%\documentclass[twocolum]{revtex4}
%%\documentclass{report}
%%\documentclass[useAMS,usenatbib, onecolumn]{mn2ejm}

@@

\begin{abstract}
% Many high-state non-magnetic cataclysmic variables (CVs) exhibit
% blue-shifted absorption or P-Cygni profiles associated with ultraviolet
% (UV) resonance lines, such as C~{\sc iv}~$\lambda1550$. These features
% imply the existence of powerful accretion disk winds in
% CVs. Here, we use an improved version of our Monte Carlo ionization
% and radiative transfer code to investigate whether disk wind models that
% produce realistic UV line profiles are also likely to generate
% observationally significant recombination line and continuum emission
% in the {\em optical} waveband. 
% We also test whether outflows may be responsible for the single-peaked 
% emission line profiles often seen
% in high-state CVs and for the weakness of the Balmer absorption edge
% (relative to simple models of optically thick accretion disks).
% We find that a standard disk wind model that is successful in
% reproducing the UV spectra of CVs also leaves a noticeable imprint on
% the optical spectrum, particularly for systems viewed at high
% inclination. The strongest optical wind-formed recombination lines are \ha\ and
% \heiiopt, although the outflow also produces emission in the Ly$\alpha$
% and \heiiuv\ UV lines. However, all of the optical lines produced by
% this default wind model are double-peaked, and the effect of the wind
% on the Balmer jump is modest. We demonstrate that a higher-density outflow model 
% displays reasonable verisimilitude with the optical spectrum of RW Trianguli,
% producing all the observed lines. This model
% produces a recombination continuum that is strong enough to fill in the Balmer jump,
% but it also predicts stronger-than-observed collisionally excited UV emission lines. 
% The wind in our revised model also alters the spectral shape of the continuum, 
% and displays a strong inclination dependence in its emission properties. 
% No single-peaked emission is seen, although we observe a 
% narrowing of the double-peaked emission lines from the base of the wind.
% We show that even denser models can produce a single-peaked \ha\ line.
% In this model, the single-peaked line profile is due to the line emission originating from 
% $\sim10^{11}$~cm above the disk plane where rotational velocities are small, 
% and is not due to the radial velocity shear mechanism that
% has previously been proposed. On the basis of our results, we suggest that winds can 
% modify, and perhaps even dominate, the line and continuum emission
% from CVs.
Many high-state non-magnetic cataclysmic variables (CVs) exhibit
blue-shifted absorption or P-Cygni profiles associated with ultraviolet
(UV) resonance lines. These features imply the existence of 
powerful accretion disk winds in CVs. Here, we use our Monte Carlo ionization
and radiative transfer code to investigate whether disk wind models that
produce realistic UV line profiles are also likely to generate
observationally significant recombination line and continuum emission
in the {\em optical} waveband. We also test whether outflows may be responsible for 
the single-peaked emission line profiles often seen in high-state CVs and for the weakness
of the Balmer absorption edge (relative to simple models of optically thick accretion disks).
We find that a standard disk wind model that is successful in
reproducing the UV spectra of CVs also leaves a noticeable imprint on
the optical spectrum, particularly for systems viewed at high
inclination. The strongest optical wind-formed recombination lines are \ha\ and
\heiiopt. We demonstrate that a higher-density outflow model produces 
all the expected H and He lines and produces a recombination continuum that can 
fill in the Balmer jump at high inclinations.
This model displays reasonable verisimilitude with the optical spectrum of RW Trianguli. No single-peaked emission is seen, although we observe a 
narrowing of the double-peaked emission lines from the base of the wind.
Finally, we show that even denser models can produce a single-peaked \ha\ line.
On the basis of our results, we suggest that winds can 
modify, and perhaps even dominate, the line and continuum emission
from CVs.
\end{abstract}

@@

\maketitle

@@

\begin{figure}	%fullpage
\centering
\includegraphics[width=0.7\textwidth]{figures/fig1.eps}
\caption{
Optical spectra of three nova-like variables: 
RW Sex (top; Beuermann et al. 1992),
IX Vel (top middle; A. F. Pala \& B. T. Gaensicke, private communication) 
and RW Tri in and out of eclipse (bottom two panels; Groot et al. 2004).
The data for RW Sex and RW Tri were digitized from the respective publications,
and the IX Vel spectrum was obtained using the XSHOOTER spectrograph 
on the Very Large Telescope on 2014 October 10.
These systems have approximate inclinations of $30^\circ$, $60^\circ$ and $80^\circ$ 
(see section 5.4) respectively. 
The trend of increasing Balmer line emission with inclination can be seen.
In RW Tri strong single-peaked emission in the Balmer lines is seen even
in eclipse, indicating that the lines may be formed in a spatially
extensive disk wind, and there is even a suggestion 
of a (potentially wind-formed) recombination continuum in the eclipsed
spectrum. We have attempted to show each spectrum over a similar dynamic range.
}
\label{novalikes}
\end{figure}

@@

\begin{figure} %fullpage
\includegraphics[width=0.9\textwidth]{figures/fig5.eps}
\caption{
The physical properties of the wind -- note the logarithmic scale. 
Near the disk plane the wind is dense, with low poloidal velocities.
As the wind accelerates it becomes less dense
and more highly ionized. The dominant He ion
is almost always He III, apart from in a small
portion of the wind at the base, which is partially shielded
from the inner disk.
}
\label{wind}
\end{figure} %fullpage

@@

\begin{figure} %fullpage
\includegraphics[width=0.9\textwidth]{figures/modela_uv_opt.eps}
\caption{
UV (left) and optical (right) synthetic spectra for model A, our benchmark model,
computed at sightlines of 10, 27.5, 45, 62.5 and 80 degrees.	
The inset plots show zoomed-in line profiles for 
\heiiuv\ and \ha. Double-peaked line emission can be seen in 
\heiiuv, \heiiopt, \ha\ and some He I lines, but the 
line emission is not always sufficient to overcome the absorption
cores from the stellar atmosphere models. The model
also produces a prominent \heiioptnew\ line at high inclinations.
}
\label{spec}
\end{figure} %fullpage\

@@

\begin{figure} %fullpage
\includegraphics[width=0.9\textwidth]{figures/modelb_uv_opt.eps}
\caption{
UV (left) and optical (right) synthetic spectra for model B computed at
sightlines of 10, 27.5, 45, 62.5 and 80 degrees. 
Model A is shown in grey for comparison.	
The inset plots show zoomed-in line profiles for 
\heiiuv\ and \ha. The Balmer and He
are double-peaked, albeit with narrower profiles.
Strong \heiiopt\ emission can be seen, as well as a trend
of a deeper Balmer jump with decreasing inclination.
}
\label{uvoptb}
\end{figure} %fullpage

@@

\begin{figure} %fullpage
\includegraphics[width=0.9\textwidth]{figures/fig13.eps}
\caption{{\sl Top Panel:} In and out of eclipse spectra of the high
inclination NL RW Tri. {\sl Bottom Panel:} In and out of eclipse synthetic
spectra from model B.
The artificial `absorption' feature just redward of the Balmer jump
is caused for the reasons described in section 5.2.}
\label{rwtricomp}
\end{figure} %fullpage

@@

>>>>

\documentclass[useAMS,usenatbib]{mn2e_x}
%%%% Various options for document class.

@@

\maketitle
\begin{abstract}
Many high-state non-magnetic cataclysmic variables (CVs) exhibit
blue-shifted absorption or P-Cygni profiles associated with ultraviolet
(UV) resonance lines. These features imply the existence of 
powerful accretion disk winds in CVs. Here, we use our Monte Carlo ionization
and radiative transfer code to investigate whether disk wind models that
produce realistic UV line profiles are also likely to generate
observationally significant recombination line and continuum emission
in the {\em optical} waveband. We also test whether outflows may be responsible for 
the single-peaked emission line profiles often seen in high-state CVs and for the weakness
of the Balmer absorption edge (relative to simple models of optically thick accretion disks).
We find that a standard disk wind model that is successful in
reproducing the UV spectra of CVs also leaves a noticeable imprint on
the optical spectrum, particularly for systems viewed at high
inclination. The strongest optical wind-formed recombination lines are \ha\ and
\heiiopt. We demonstrate that a higher-density outflow model produces 
all the expected H and He lines and produces a recombination continuum that can 
fill in the Balmer jump at high inclinations.
This model displays reasonable verisimilitude with the optical spectrum of RW Trianguli. No single-peaked emission is seen, although we observe a 
narrowing of the double-peaked emission lines from the base of the wind.
Finally, we show that even denser models can produce a single-peaked \ha\ line.
On the basis of our results, we suggest that winds can 
modify, and perhaps even dominate, the line and continuum emission
from CVs.
\end{abstract}

@@

@@

\begin{figure*}
\centering
\includegraphics[width=0.7\textwidth]{figures/fig1.eps}
\caption{
Optical spectra of three nova-like variables: 
RW Sex (top; Beuermann et al. 1992),
IX Vel (top middle; A. F. Pala \& B. T. Gaensicke, private communication) 
and RW Tri in and out of eclipse (bottom two panels; Groot et al. 2004).
The data for RW Sex and RW Tri were digitized from the respective publications,
and the IX Vel spectrum was obtained using the XSHOOTER spectrograph 
on the Very Large Telescope on 2014 October 10.
These systems have approximate inclinations of $30^\circ$, $60^\circ$ and $80^\circ$ 
(see section 5.4) respectively. 
The trend of increasing Balmer line emission with inclination can be seen.
In RW Tri strong single-peaked emission in the Balmer lines is seen even
in eclipse, indicating that the lines may be formed in a spatially
extensive disk wind, and there is even a suggestion 
of a (potentially wind-formed) recombination continuum in the eclipsed
spectrum. We have attempted to show each spectrum over a similar dynamic range.
}
\label{novalikes}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=0.8\textwidth]{figures/fig5.eps}
\caption{
The physical properties of the wind -- note the logarithmic scale. 
Near the disk plane the wind is dense, with low poloidal velocities.
As the wind accelerates it becomes less dense
and more highly ionized. The dominant He ion
is almost always He III, apart from in a small
portion of the wind at the base, which is partially shielded
from the inner disk.
}
\label{wind}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=0.9\textwidth]{figures/modela_uv_opt.eps}
\caption{
UV (left) and optical (right) synthetic spectra for model A, our benchmark model,
computed at sightlines of 10, 27.5, 45, 62.5 and 80 degrees.	
The inset plots show zoomed-in line profiles for 
\heiiuv\ and \ha. Double-peaked line emission can be seen in 
\heiiuv, \heiiopt, \ha\ and some He I lines, but the 
line emission is not always sufficient to overcome the absorption
cores from the stellar atmosphere models. The model
also produces a prominent \heiioptnew\ line at high inclinations.
}
\label{spec}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=0.9\textwidth]{figures/modelb_uv_opt.eps}
\caption{
UV (left) and optical (right) synthetic spectra for model B computed at
sightlines of 10, 27.5, 45, 62.5 and 80 degrees. 
Model A is shown in grey for comparison.	
The inset plots show zoomed-in line profiles for 
\heiiuv\ and \ha. The Balmer and He
are double-peaked, albeit with narrower profiles.
Strong \heiiopt\ emission can be seen, as well as a trend
of a deeper Balmer jump with decreasing inclination.
}
\label{uvoptb}
\end{figure*}

@@

@@

\begin{figure*}
\includegraphics[width=0.9\textwidth]{figures/fig13.eps}
\caption{{\sl Top Panel:} In and out of eclipse spectra of the high
inclination NL RW Tri. {\sl Bottom Panel:} In and out of eclipse synthetic
spectra from model B.
The artificial `absorption' feature just redward of the Balmer jump
is caused for the reasons described in section 5.2.}
\label{rwtricomp}
\end{figure*}
